---
title: "FCC and ACS Pre-processing and EDA"
output: html_document
---

The following document analyzes complaints filed to the Federal Communications Commission (FCC) from January 2021 to July 2022. This data set was pulled from the FCC's website. To better understand the demographics of the individuals who filed complaints, the FCC data set was merged with a subset of data from the American Community Survey (ACS) which comes from the United States Census
Bureau. _Analyzing complaints about spam phone calls was an area of interest._

__Data Sources:__

* fcc_complaints_CA-2021.csv
* acs_data.csv

__Below is a list of variables that were of importance:__

FCC Data:

* zip - the zip code where the complaint originated from.
* caller_id_number - the phone number from which the spam or robocall came from.
* date_of_issue - the date the incident occurred.
* time_of_issue - the time of day the incident occurred.
* method - how the spam or robocall connected with the individual.

ACS Data:

* name - contains the ZIP code areas.
* median_income - contains the median income in each ZIP code.
* median_age - contains the median age in each ZIP code.
* total_pop - contains the total population within each ZIP code.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

#### Loading and cleaning column names of data

```{r}
fcc <- janitor::clean_names(read_csv('fcc_complaints_CA-2021.csv'))
```

#### Filtering out n/a values in zip & method columns

```{r}
fcc_clean <- fcc %>%
  drop_na(method, zip) %>%
  filter(zip != '00000')
```

#### Creating a new column that extracts tbe area code from caller_id column

```{r}
fcc_clean <- fcc_clean %>%
  extract(
    col = caller_id_number,
    into = 'area_code',
    regex = "(\\d\\d\\d)",
    remove = F
  )
```

#### Creating separate columns for month and year of entry

```{r}
fcc_clean$date_of_issue <- lubridate::ymd(fcc_clean$date_of_issue)
fcc_clean <- fcc_clean %>%
  mutate(
    month_of_issue = lubridate::month(date_of_issue, label = T),
    year_of_issue = lubridate::year(date_of_issue)
  )
  
```

#### Formatting and standardizing time values in column

```{r}
fcc_clean$time_of_issue <-
  str_replace_all(fcc_clean$time_of_issue, "\\.", '')
fcc_clean <- fcc_clean %>%
  mutate(time_of_issue_clean = parse_time(fcc_clean$time_of_issue))
```

#### Visualizing the distribution of phone complaints by method 

```{r}
call_distribution <- fcc_clean %>%
  filter(form == "Phone") %>%
  ggplot(aes(time_of_issue_clean, fill = method)) + ## edit the fill
  geom_histogram(alpha = 0.8) +
  theme_minimal() +
  scale_x_time(
    breaks = scales::date_breaks("3 hour"),
    labels = scales::time_format("%I:%M %p")
  ) + ## edit the time_format
  theme(legend.position = "bottom") + ## edit this line
  labs(
    x = "",
    y = "Number of FCC Complaints",
    fill = "",
    title = "Distribution of Phone Complaints to the FCC",
    subtitle = "From January 2021 - July 2022"
  ) +
  ggthemes::scale_fill_fivethirtyeight()
```

```{r, echo = FALSE}
plot(call_distribution)
```

#### Reading acs data

```{r}
acs <- read_csv('acs_data.csv')
```

#### Creating new column of zip code values

```{r}
acs <- acs %>%
  extract(
    col = name,
    into = 'zip',
    regex = "(\\d\\d\\d\\d\\d)",
    remove = F
  )
```

#### Left joining acs with fcc_clean by zip codes and filtering year to be 2021 and population to not equal 0

```{r}
fcc_joined <- fcc_clean %>%
  merge(acs, by = 'zip') %>%
  filter(year_of_issue == 2021,
         total_pop != 0)
```

#### Creating columns to easily distinguish income and age status of individuals

```{r}
fcc_joined <- fcc_joined %>%
  mutate(
    low_income_zip = ifelse(
      median_income < mean(median_income, na.rm = T),
      'Below Average Median Income',
      'Above Average Median Income'
    ),
    high_age_zip = ifelse(
      median_age > mean(median_age, na.rm = T),
      'Above Average Median Age',
      'Below Average Median Age'
    )
  )

```

#### Creating a tibble that only includes phone complaints and a tibble that only includes spam complaints

```{r}
fcc_joined_spam <- fcc_joined %>%
  filter(form == 'Phone' & issue == 'Unwanted Calls')

fcc_joined_phone <- fcc_joined %>%
  filter(form == 'Phone')

```

#### Summarizing the merged tibble to count the numbers of demographic similarities

```{r}
aggregate <- fcc_joined %>%
  group_by(median_age, median_income, total_pop, zip, low_income_zip) %>%
  summarise(n_complaints = n())
```
```{r, echo=FALSE}
show(aggregate)
```

#### Creating a column that calculates the proportion of complaints per-1000 to adjust for differences in population between zip codes

```{r}
aggregate <- aggregate %>%
  mutate(complaints_per_1000 = n_complaints/total_pop*1000)
```

#### Adding average median income and average median age to spam call dataset

```{r}
fcc_joined_spam <- fcc_joined_spam %>%
  mutate(
    low_income_zip =  mean(median_income, na.rm = T),
    high_age_zip =mean(median_age, na.rm = T)
  )
```

#### Counting the number of spam call complaints by zip code

```{r}
dominant_city <- head(count(fcc_joined_spam, zip, sort = T))
```
```{r, echo=FALSE}
show(dominant_city)
```


#### Counting the number of spam call complaints by month

```{r}
dominant_time_of_year <- head(count(fcc_joined_spam, month_of_issue, sort = T))
```
```{r,echo=FALSE}
show(dominant_time_of_year)
```

#### Counting the number of spam call complaints to the FCC of the dominant income level

```{r}
dominant_income <- count(fcc_joined_spam, low_income_zip,sort = T) %>%
  drop_na()
colnames(dominant_income) <- c('Income Level', 'Number of Complaints to the FCC')
```
```{r, echo=FALSE}
show(dominant_income)
```


#### Counting the number of spam call complaints to the FCC by the dominant age level

```{r}
dominant_age <- count(fcc_joined_spam, high_age_zip, sort = T) %>%
  drop_na()
colnames(dominant_age) <- c('Age Level', 'Number of Complaints to the FCC')
```
```{r, echo=FALSE}
show(dominant_age)
```


#### Finding average median income of those that complained about spam calls

```{r}
average_income <- mean(fcc_joined_spam$median_income, na.rm = T)
```
```{r, echo=FALSE}
show(average_income)
```


#### Visualizing the number of spam call complaints to the FCC by zip-codes

```{r}
dominant_cities <-head(dominant_city, 10) %>%
  ggplot(aes(x = reorder(zip, -n), y = n)) +
  geom_bar(stat = 'identity', fill="lightblue") +
  labs(x = 'Zip code', y = 'Number of complaints to the FCC', title = "Zip Codes With the Most Complaints to the FCC ",subtitle = '(Only In 2021)') +
  scale_fill_manual(values=c("lightblue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 

```

```{r, echo = FALSE}
plot(dominant_cities)
```

#### Visualizing the number of spam call complaints to the FCC by the month of the year

```{r}
time_of_year_calls <- dominant_time_of_year %>%
  ggplot(aes(x = month_of_issue, y = n, group = 1)) +
  labs(x = 'Month of the year', y = 'Number of complaints to the FCC', title = 'Number of Complaints to the FCC by Month',subtitle = '(Only In 2021)')+
  geom_line() +
  geom_point()+
  theme_minimal()
```

```{r, echo = FALSE}
plot(time_of_year_calls)
```

#### Finding the average median income by zip-code

```{r}
income_by_zip <- fcc_joined_spam %>%
  group_by(zip) %>%
  summarise(average_median_income = round(mean(median_income, na.rm = T)))
```
```{r, echo=FALSE}
show(income_by_zip)
```

#### Merging the data set of the zip-codes with the most complaints and the data set of the average median income of the zip-codes  

```{r}
dominant_city <- merge(dominant_city, income_by_zip, by = 'zip')
dominant_city <- dominant_city[order(dominant_city$n, decreasing = TRUE),]
```

#### Visualizing the number of complaints to FCC by zip-codes and overlaying the respective zip-codes' average median income

```{r}
dominant_cities_income_graph <-head(dominant_city, 10) %>%
  ggplot(aes(x = reorder(zip, -n), y = n)) +
  geom_bar(stat = 'identity', fill="lightblue") +
  labs(x = 'Zip code', y = 'Number of complaints to the FCC', title = "Zip Codes With the Most Complaints to the FCC",subtitle = '(Only In 2021)') +
  scale_fill_manual(values=c("lightblue")) +
  theme_minimal() +
  geom_text(aes(label=average_median_income, vjust=-0.5, colour = 'average median income '))+
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```

```{r, echo = FALSE}
plot(dominant_cities_income_graph)
```

#### Finding the percentage of phone related complaints to the FCC that were about spam calls

```{r}
spam_percentage <- nrow(fcc_joined_spam)/nrow(fcc_joined_phone)
```
```{r, echo=FALSE}
show(spam_percentage)
```

#### Counting the number of complaints to FCC by zip-code

```{r}
zip_density <- fcc_joined_spam %>%
  count(zip, sort = T)
colnames(zip_density) = c('region', 'value')
```

#### Loading in library for choropleth

```{r}
library(choroplethrZip)
```

#### Creating choropleth of complaints to FCC by zip-codes in California

```{r}
zip_choropleth(zip_density, title = "Density Map of FCC Complaints by Zip Codes", legend = "Number of Complaints", num_colors = 6,
               state_zoom = 'california')
```
